---
title: "Data Analisys part 2"
subtitle: "lession 4c lubridate"
author: "Eviatar Guttman"
date: "Last updated: `r format(Sys.time(), '%d-%m-%Y')`"  
output: 
  html_document: 
    toc: yes
    toc_depth: 3
    toc_float: yes
    number_sections: no
    toc_collapsed: no
    smooth_scroll: yes
    fig_caption: yes
    highlight: tango
    theme: flatly
    css: ../style.css
urlcolor: blue
editor_options: 
  chunk_output_type: console
---  
```{r setup, include = F}
# global chunk options
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      tidy=TRUE,
                      # fig.path='Figures_RMD/',
                      fig.align = 'center',
                      collapse = TRUE)

# don't show NA's in tables
options(knitr.kable.NA = '')

```

```{r}
library(tidyverse)
library(nycflights13)
library(lubridate)
```

There are three types of date/time data that refer to an instant in time:  
    
1. A date.  Tibbles print this as `<date>`.  
2. A time within a day.  Tibbles print this as `<time>`.  
3. A date-time is a date plus a time: it uniquely identifies an instant in time (typically to the nearest second). Tibbles print this as `<dttm>`. Elsewhere in R these are called POSIXct.  

```{r}
today()

attributes(today())

typeof(today())


now()

attributes(now())

typeof(now())


```

There are three ways you’re likely to create a date/time:  
    
1. From a string.  
2. From individual date-time components.  
3. From an existing date/time object.  

# from strings 
Here are a banch of functions that Transforms dates stored in character and numeric vectors to Date or POSIXct objects: 
```{r}
ymd("2017-01-31")

ymd("2017/01/31")

mdy("January 31st, 2017")

dmy("31-Jan-2017")

x <- c(20090101, "2009-01-02", "2009 01 03", "2009-1-4",
       "2009-1, 5", "Created on 2009 1 6", "200901 !!! 07")
ymd(x)

ymd_hms("2017-01-31 20:11:59")

mdy_hm("01/31/2017 08:01")
```


# From individual components  
`make_date()` produces objects of class `Date`
```{r}
make_date(year = 1970L, month = 1L, day = 1L)
```

like most functions in R, `make_date` is vectorized: 
```{r}
flights %>% 
  select(year, month, day) %>% 
  mutate(departure = make_date(year, month, day))

flights %>% 
  select(year, month, day, hour, minute) %>% 
  mutate(departure = make_datetime(year, month, day, hour, minute))

```

# From other types  
You may want to switch between a date-time and a date. That’s the job of `as_datetime()` and `as_date()`:

```{r}
as_datetime(today())

as_date(now())
```

 Sometimes you’ll get date/times as numeric offsets from the “Unix Epoch”, 1970-01-01.  
 If the offset is in seconds, use `as_datetime()`; if it’s in days, use `as_date()`.
```{r}
60 * 60 * 10 # 10 hours

as_datetime(60 * 60 * 10)

365 * 10 + 2 # 10 years

as_date(365 * 10 + 2)
```

# Useful functions

```{r}
some_time <- ymd_hms("2016-07-08 12:34:56")

year(some_time)

month(some_time)

month(some_time, label = T)

month(some_time, label = T, abbr = F)
```

This happens because we set: `Sys.setlocale("LC_ALL", "hebrew")` (I think).
there is a function called `local()` that reads the days and time from you system locality. 

```{r}
locale()

locale('he')
```

count number of X since the beginning

```{r}
mday(some_time) # מספר ימים מתחילת החודש

yday(some_time) # מספר ימים מתחילת שנה

wday(some_time) # יום שישי

wday(some_time, label = T)
```

its an ordered factor. we haven't talked about it yet. 
```{r}
attributes(wday(some_time, label = T))
```

We can use wday() to see that more flights depart during the week than on the weekend  
```{r}
names(flights)

flights %>% select(year, month, day, dep_time  ) %>%
  mutate(departure = make_datetime(year, month, day, dep_time %/% 100, dep_time %% 100)) %>%
  mutate(wday = wday(departure, label = T)) %>%
  ggplot(aes(x = wday))+ geom_bar()

```
we need to check about NA's in the data!

Very useful is : `floor_date`, `round_date`, `ceiling_date`

```{r}
(x <- ymd_hms("2009-08-03 12:01:59.23"))

round_date(x, "second")

round_date(x, "minute")

round_date(x, "5 mins")

round_date(x, "2 hours")

round_date(x, "bimonth")

round_date(x, "quarter") == round_date(x, "3 months")

round_date(x, "halfyear")

round_date(x, "year")


```

# difftime
there is also a timediff format:   


```{r}
arrive <- ymd_hms("2011-06-04 12:00:00", tz = "Pacific/Auckland")

leave <- ymd_hms("2011-08-10 14:00:00", tz = "Pacific/Auckland")

(time_diff <- interval(arrive, leave))

attributes(time_diff)
```
number of days differnce: 

```{r}
time_diff / ddays(1)

time_diff / ddays(3)

time_diff / dweeks()
```

number of month diff: 

```{r}
time_diff / dmonths()
```


























<div dir="rtl">

</div>


# exercises  
<div dir="rtl">
1. מה קורה אם מנסים לעשות parse ל-string שמכיל תאריכים לא תקפים?   
</div>
`ymd(c("2010-10-10", "bananas"))`  
<div dir="rtl">
2. מה התפקיד של הארגומנט `tzone` בפונקציה `today`? מדוע הוא חשוב?   
3. השתמש בפונקציות המתאימות ב-`lubridate` כדי לעשות parse לתאריכים הבאים: 
</div>
```{r}
d1 <- "January 1, 2010"
d2 <- "2015-Mar-07"
d3 <- "06-Jun-2017"
d4 <- c("August 19 (2015)", "July 1 (2015)")
d5 <- "12/30/14" # Dec 30, 2014
```

<div dir="rtl">
4. איך ההתפלגות של שעות הטיסה משתנות לאורך השנה?   
5. באיזה יום בשבוע כדאי לטוס אם רוצים למזער את ההסתברות לאיחור?   
</div>


